---
title: "forecasting"
author: "Thi Allgood"
date: "9/11/2020"
output: html_document
---
---
title: "Forecasting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r }
library(odbc)
library(tidyverse)
library(odbc)
library(forecast)
library(RcppRoll)
library(tsibble)
library(tsibbledata)
library(ggplot2)
library(feasts)
library(ggplot2)
library(timetk)
library(zoo)
library(GGally)
library(readxl)
library(fpp2)
library(fpp3)
```



```{r }
con <- dbConnect(odbc::odbc(), "CLPImpala")
depletions_df <- dbGetQuery(con, 
           "SELECT brand_name,
           t_date, 
           sum(depl_9l) as 'depletions'
           FROM dp_data_science.vwb_volumetric_depletions
           WHERE mkt_lvl3 = 'United States'
           AND brand_lvl1 = 'Active Brands'
           AND value_type = '010'
           GROUP BY t_date, brand_name
           ORDER BY t_date") 

#List of brands to be forecast
```

```{r}
brand_list <- list("EL JIMADOR",
                   "HERRADURA", 
                   "JACK DANIELS TENNESS")
brand = brand_list[[3]]
```

Brands
```{r}
depletions_df1 <- depletions_df %>%
  group_by(brand_name, t_date) %>%
  summarize(depletions = mean(depletions)) %>%
  mutate(t_date = yearmonth(t_date)) %>%
  filter(brand_name %in% brand_list)

depletions_df1 <- depletions_df1 %>%
  slice(1:(nrow(depletions_df1)-2))



tsibble1 <- depletions_df1 %>%
  as_tsibble(key=brand_name,index = t_date)

```

```{r}
tsibble1 %>%
  autoplot()

tsibble1 %>%
  tsibble::fill_gaps() %>%
  gg_season(depletions,period = "year")
```


Brand: JACK DANIELS TENNESS
```{r}
brand_df <- depletions_df1 %>%
  filter(brand_name == brand) %>%
  mutate(t_date = yearmonth(t_date))

brand_df <- brand_df %>%
  slice(1:(nrow(brand_df)-2))

tsibble <- brand_df %>%
  as_tsibble(key=brand_name,index = t_date)

tsibble %>%
  autoplot() +
  ylab("JACK DANIELS TENNESS") + xlab("Month")

tsibble %>%
  tsibble::fill_gaps() %>%
  gg_season(depletions,period = "year") + 
  ylab("JACK DANIELS TENNESS") + xlab("Month")

tsibble %>%
  gg_subseries(depletions) +
  ylab("JACK DANIELS TENNESS")

tsibble %>%
  gg_lag(depletions, geom="point")

tsibble %>%
  autoplot()

tsibble %>% ACF(depletions) %>%
  autoplot()

```


```{r}
dcmp <- tsibble %>%
  model(STL(depletions))
components(dcmp)
```

```{r}
tsibble %>%
  autoplot(depletions,color = "gray") +
  autolayer(components(dcmp), trend, color = "red") + 
  xlab("Month") + ylab("Depletions")
```


```{r}
tsibble %>%
  model(classical_decomposition(depletions,type = 'additive')) %>%
  components() %>%
  autoplot() + xlab("Month")
```


```{r}
tsibble %>%
  features(depletions,mean)

tsibble %>%
  features(depletions, quantile, prob=seq(0,1,by=0.25))

tsibble %>%
  features(depletions,feat_acf)

tsibble %>%
  features(depletions,feat_stl)

tsibble %>%
  features(depletions, feat_stl) %>%
  ggplot(aes(x= trend_strength, y = seasonal_strength_year, col= brand_name)) +
  geom_point()


```

# The forecaster's toolbox

## Model: Time series linear model

```{r}
fit <- tsibble %>%
  model(trend_model = TSLM(depletions ~ trend()))
fit %>%
  forecast(h = "3 years") %>%
  autoplot(tsibble) 
```

## Model: average method

```{r}
fit2 <- tsibble %>%
  model(MEAN(depletions)) 

fit2 %>%
  forecast(h = "3 years") %>%
  autoplot(tsibble)
```

## Model: naive

```{r}
fit3 <- tsibble %>%
  model(NAIVE(depletions)) 

fit3 %>%
  forecast(h = "3 years") %>%
  autoplot(tsibble)
```

## Model: seasonal naive method

```{r}
fit4 <- tsibble %>%
  model(SNAIVE(depletions)) 

fit4 %>%
  forecast(h = "3 years") %>%
  autoplot(tsibble)
```

## Model: drift method

```{r}
fit5 <- tsibble %>%
  model(RW(depletions ~ drift())) 

fit5 %>%
  forecast(h = "3 years") %>%
  autoplot(tsibble)
```

# Residual diagnostics

```{r}
aug <- tsibble %>%
  model(NAIVE(depletions)) %>%
  augment()

aug %>%
  autoplot(.resid) +
  xlab("Depletions") +
  ggtitle("Residuals from naïve method")
```


```{r}
aug %>%
  ggplot(aes(x = .resid)) +
  geom_histogram() +
  ggtitle("Histogram of residuals")
```

```{r}
aug %>%
  ACF(.resid) %>%
  autoplot() + 
  ggtitle("ACF of residuals")

```


```{r}
tsibble %>% model(NAIVE(depletions)) %>% gg_tsresiduals()
```

#  Evaluating forecast accuracy
## Filter

```{r}
tsibble %>%
  filter(year(t_date) >= 2019)
tsibble %>%
  filter(quarter(t_date) == 1)
tsibble %>%
  slice(n()-11:0)
```

```{r fig1, fig.height = 20, fig.width = 40}

tsibble.train <- tsibble %>%
  filter(year(t_date)<2019)

fit6 <- tsibble.train %>%
  model(
    Mean = MEAN(depletions),
    Naive = NAIVE(depletions),
    Seasonal.naive = SNAIVE(depletions),
    Drifts = RW(depletions ~ drift())
  )

tsibble.forecast <- fit6 %>%
  forecast(h=19)

tsibble.forecast %>%
  autoplot(tsibble, level = NULL) 
```

```{r}
accuracy(tsibble.forecast, tsibble)
```
## Time series cross-validation

```{r}
tsibble.tr <- tsibble %>%
  slice(1:(n()-1)) %>%
  stretch_tsibble(.init = 3, .step = 1)

tsibble.forecast2 <- tsibble.tr %>%
  model(RW(depletions ~ drift())) %>%
  forecast(h=1)

tsibble.forecast2 %>%
  accuracy(tsibble)

tsibble %>%
  model(RW(depletions ~ drift())) %>%
  accuracy()
```



```{r}
tsibble.tr.12 <- tsibble %>%
  slice(1:(n()-12)) %>%
  stretch_tsibble(.init = 3, .step = 1)

fit7 <- tsibble.tr.12 %>%
  model(RW(depletions ~ drift())) %>%
  forecast(h=12) %>%
  group_by(.id) %>%
  mutate(h=row_number()) %>%
  ungroup()
fit7 %>%
  accuracy(tsibble,by="h") %>%
  ggplot(aes(x=h,y=RMSE)) + geom_point()
```

# Simple exponential smoothing
 
```{r}

fit8 <- tsibble %>%
  model(ETS(depletions ~ error("A") + trend("N") + season("N"), opt_crit = "mse"))
fc <- fit8 %>%
  forecast(h = 5)
```

```{r}
fc %>%
  autoplot(tsibble) +
  geom_line(aes(y=.fitted, colour = "Fitted"), data=augment(fit8)) +
  xlab("Month")
```


```{r}
tsibble %>%
  model(
    `Holt's method` = ETS(depletions ~ error("A") + trend("A") + season("N")),
    `Damped Holt's method` = ETS(depletions ~ error("A") + trend("Ad", phi = 0.9) + season("N"))
  ) %>%
  forecast(h=15) %>%
  autoplot(tsibble,level = NULL) 


```

## Method with seasonality

```{r}
fit9 <- tsibble %>%
  summarise(depletions = sum(depletions)) %>%
  model(
    additive = ETS(depletions ~ error("A") + trend("A") + season("A")),
    multiplicative = ETS(depletions ~ error("M") + trend("A") + season("M"))
  )

forecast1 <- fit9 %>%
  forecast(h="3 years")

forecast1 %>%
  autoplot(tsibble, level = NULL) + xlab("Month") +
  scale_color_brewer(type = "qual", palette = "Dark2")

```

## Estimating ETS models

```{r}
fit10 <- tsibble %>%
  summarise(depletions = sum(depletions)) %>%
  model(ETS(depletions))

report(fit10)
```


```{r}
components(fit10) %>%
  autoplot() +
  ggtitle("ETS(M,N,M) components")
```



```{r}
residuals(fit10) %>%
  autoplot()
residuals(fit10, type = "response") %>%
  autoplot()
```

## Forecasting with ETS models



```{r fig3, fig.height = 10, fig.width = 15 }
fit10 %>%
  forecast(h = 36) %>%
  autoplot(tsibble)

```

# ARIMA
## Seasonal differencing



```{r }

tsibble %>%
  summarise(depletions = sum(depletions)) %>%
  transmute(
    depletions = depletions,
    Log.depletion = log(depletions),
    annual.change.in.log = difference(log(depletions), 12),
    Doubly.difference.log = difference(difference(log(depletions), 12), 1)
  ) %>%
  gather("Type","depletions", !!!syms(measured_vars(.)), factor_key = T) %>%
  ggplot(aes(x = t_date, y = depletions)) +
  geom_line() + 
  facet_grid(vars(Type), scale = "free_y")

```

## Non seasonal arima models
```{r}
arima.fit <- tsibble %>%
  model(ARIMA(depletions ~ PDQ(0,0,0)))
report(arima.fit)
arima.fit %>%
  forecast(h=24) %>%
  autoplot(slice(tsibble, (n()-80):n()))


```



### Seasonality
#### ACF 

```{r}
tsibble %>%
  ACF(depletions) %>%
  autoplot()
```

#### PACF

```{r}
#
tsibble %>%
  PACF(depletions) %>%
  autoplot()

```

```{r}
#
fit11 <- tsibble %>%
  model(ARIMA(depletions ~ pdq(1,0,0) + PDQ(0,0,0)))
report(fit11)
```

### Seasonally adjusted ?

```{r}
tsibble.seasonally.adjusted <- tsibble %>%
  model(STL(depletions ~ season(window = "periodic"))) %>%
  components() %>%
  select(-.model) %>%
  as_tsibble()
tsibble %>%
  autoplot()
tsibble.seasonally.adjusted %>%
  autoplot(season_adjust)


```

```{r}
tsibble.seasonally.adjusted %>%
  gg_tsdisplay(difference(season_adjust), plot_type = 'partial')


```


```{r}
# arima110
fit110 <- tsibble.seasonally.adjusted %>%
  model(
    arima = ARIMA(season_adjust ~ pdq(1,1,0) + PDQ(0,0,0))
  )
report(fit110)

```

```{r}
# arima411
fit.arima <- tsibble.seasonally.adjusted %>%
  model(
    arima = ARIMA(season_adjust)
  )
report(fit.arima)


```

```{r}
#
fit.arima %>%
   gg_tsresiduals()

```

```{r}
#
fit.arima %>% forecast() %>% autoplot(tsibble.seasonally.adjusted)


```


```{r}
gg_arma(fit.arima)


```
##  Seasonal ARIMA models

```{r}
#
tsibble %>% 
  gg_tsdisplay(depletions %>% difference(12) %>% difference(), plot_type='partial') 

```


```{r}
#


```

```{r}
#


```

```{r}
#


```

